---
title: "Group 10: Spotify Data Analysis Final"
author: "Kassidy, Lejla, Maya"
date: "2024-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library(tidyr)
``` 

```{r}
spotify2023 <- read_csv("top_50_2023.csv") %>% clean_names()
```

```{r}
split_spotify2023 <- spotify2023 %>% 
  separate(genres, into = c("genre1", "genre2", "genre3", "genre4", "genre5", "genre6"), sep=",") %>% 
  pivot_longer(cols = starts_with("genre"),
               names_to = "type",
               values_to = "class") %>% 
  filter(class != "NA") %>% 
  select(-type)
```

## Re-use select to pull out the punctuation
```{r}
split_spotify2023 <- split_spotify2023 %>% 
  separate(class, into = c("punctuation", "class2"), sep = "'") %>% 
  select(-punctuation)
```

## pivot_wider 
```{r}
split_spotify2023 <- split_spotify2023 %>%
    pivot_wider(
        names_from = class2,
        values_from = class2,
        values_fn = length) %>% 
  mutate(across(everything(), ~replace(.x, is.na(.x), 0)))
```

## Seperate the album release data into year, month, and day
```{r}
spotify <- split_spotify2023 %>% 
  separate(album_release_date, into = c("year", "month", "day"), sep = "-") %>% 
  clean_names()
```

```{r}
names(spotify)
```


## Analyze all the songs to see which genre has the highest number of songs
```{r}
spotify %>% 
  select(pop:musica_chihuahuense) %>% 
  gather(key = "genre", value = "count") %>% 
  ggplot(aes(x = reorder(genre, count), y = count)) +
  geom_bar(stat = "identity", fill = "springgreen3", color="black") +
  labs(title = "Number of Songs in Each Genre",
       x = "Genre",
       y = "Number of Songs")+
  coord_flip()+
  theme_minimal()
```

## List all the pop songs
```{r}
spotify %>% 
  filter(pop == 1) %>% 
  select(track_name, artist_name)
```

## List all the pop songs and group by artists
```{r}
spotify %>% 
  filter(pop == 1) %>% 
  group_by(artist_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

## Total number of songs from the top 6 artists
```{r}
spotify %>%
  count(artist_name) %>%
  top_n(6) %>%
  ggplot(aes(x = reorder(artist_name, -n), y = n)) +
  geom_bar(stat = "identity", fill = "springgreen3", color="black") +
  labs(title = "Top Artists with the Most Songs",
       x = NULL,
       y = "Number of Songs")+
  theme_minimal()
```

## Analyze all the pop genres to see which one is the highest number of songs
```{r}
spotify %>% 
  select(contains("pop")) %>% 
  select(-popularity) %>% 
  gather(key = "pop", value = "count") %>% 
  ggplot(aes(x = reorder(pop, count), y = count)) +
  geom_bar(stat = "identity", fill = "springgreen3", color="black") +
  labs(title = "All Pop Songs",
       x = NULL,
       y = "Number of Songs")+
  coord_flip()+
  theme_minimal()
```

## List all the canadian_pop songs
```{r}
spotify %>% 
  filter(canadian_pop == 1) %>% 
  select(track_name, artist_name)
```

## What are all the k_pop and k_pop_girl_group songs
```{r}
spotify %>%
  filter(k_pop > 0 | k_pop_girl_group >0) %>% 
  select(artist_name, track_name) 
```

## Density graph on the duration of all songs
```{r}
spotify %>% 
  mutate(duration_ms = duration_ms/60000) %>%
  ggplot(aes(x = duration_ms)) +
  geom_density(fill = "springgreen3") +
  labs(title = " Average Duration of All Songs",
       x = "Duration (min)",
       y = "Density")+
  theme_minimal()
```

## Explicit and Non-Explicit Songs compared to loudness
```{r}
spotify %>%
  mutate(is_explicit = ifelse(is_explicit == 1, "Explicit", "Not Explicit")) %>%
  ggplot(aes(x = is_explicit, y = loudness, fill = is_explicit)) +
  geom_boxplot(fill="springgreen3") +
  labs(title = "Loudness of Explicit and Non-Explicit Songs",
       x = NULL,
       y = "Loudness")+
  theme_minimal()
```

## Dancability and energy of all the songs
```{r}
spotify %>% 
  ggplot(aes(x = danceability, y = energy)) +
  geom_point(size = 2.5) +
   geom_smooth(method = "lm", color="springgreen3") +
  labs(title = "Analysis of the Danceability and Energy of All Songs",
       x = "Danceability",
       y = "Energy")+
  theme_minimal()
```

## Dancability and tempo of all songs
```{r}
spotify %>%
  ggplot(aes(x = danceability, y = tempo)) +
  geom_point(size=2.5) +
  labs(title = "Analysis of the Danceability and Tempo of All Songs",
       x = "Danceability",
       y = "Tempo")+
  theme_minimal()
```

## Energy and tempo of all songs
```{r}
spotify %>%
  ggplot(aes(x = energy, y = tempo)) +
  geom_point(size=2.5) +
  labs(title = "Analysis of the Energy and Tempo of All Songs",
       x = "Energy",
       y = "Tempo")+
  theme_minimal()
```

## Acoustics and liveness of all songs
```{r}
spotify %>%
  ggplot(aes(x = acousticness, y = liveness)) +
  geom_point(size=2.5) +
  labs(title = "Analysis of the Acoustics and Liveness of All Songs",
       x = "Acousticness",
       y = "Liveness")+
  theme_minimal()
```

## Make a graph of the dancability and valance. Add a regression line
```{r}
spotify %>%
  ggplot(aes(x = danceability, y = valence)) +
  geom_point(size=2.5) +
  geom_smooth(method = "lm", color="springgreen3") +
  labs(title = "Analysis of the Dancability and Valance of All Songs",
       x = "Danceability",
       y = "Valence")+
  theme_minimal()
```


## Make a graph of the energy and valance. Add a regression line
```{r}
spotify %>%
  ggplot(aes(x = energy, y = valence)) +
  geom_point(size=2.5) +
  geom_smooth(method = "lm", color="springgreen3") +
  labs(title = "Analysis of the Energy and Valance of All Songs",
       x = "Energy",
       y = "Valence")+
  theme_minimal()
```



## Make a shiny app for all the graphs used so far, seperate then into different tabs on the sidebar. color the dashboardpage springgreen3
```{r} 
library(shiny)
library(shinydashboard)

ui <- dashboardPage(skin = "green",
  dashboardHeader(title = "Spotify Data Analysis", titleWidth = 350),
  dashboardSidebar(
    sidebarMenu(
      menuItem("Songs in Each Genre", tabName = "genre"),
      menuItem("Pop Songs Table", tabName = "pop"),
      menuItem("Top Artists", tabName = "top5"),
      menuItem("Pop Genres", tabName = "pop_genre"),
      menuItem("Canadian Pop Table", tabName = "canadian_pop"),
      menuItem("K-Pop and K-Pop Girl Group Table", tabName = "kpop"),
      menuItem("Average Duration", tabName = "duration"),
      menuItem("Explicit vs. Non-Explicit Songs", tabName = "loudness"),
      menuItem("Danceability and Tempo", tabName = "dance_tempo"),
      menuItem("Energy and Tempo", tabName = "energy_tempo"),
      menuItem("Acoustics and Liveness", tabName = "acoustic_liveness"),
      menuItem("Danceability and Energy", tabName = "dance_energy"),
      menuItem("Danceability and Valance", tabName = "dance_valance"),
      menuItem("Energy and Valance", tabName = "energy_valance")
    )
  ),
  dashboardBody(
    tabItems(
      tabItem(tabName = "genre",
              fluidRow(
                box(plotOutput("genre"), width = 12)
              )
      ),
      
      tabItem(tabName = "pop",
              fluidRow(
                box(tableOutput("pop"), width = 12)
              )
      ),
      tabItem(tabName = "top5",
              fluidRow(
                box(plotOutput("top5"), width = 12)
              )
      ),
      tabItem(tabName = "pop_genre",
              fluidRow(
                box(plotOutput("pop_genre"), width = 12)
              )
      ),
      tabItem(tabName = "canadian_pop",
              fluidRow(
                box(tableOutput("canadian_pop"), width = 12)
              )
      ),
      tabItem(tabName = "kpop",
              fluidRow(
                box(tableOutput("kpop"), width = 12)
              )
      ),
      tabItem(tabName = "duration",
              fluidRow(
                box(plotOutput("duration"), width = 12)
              )
      ),
      tabItem(tabName = "loudness",
              fluidRow(
                box(plotOutput("loudness"), width = 12)
              )
      ),

      tabItem(tabName = "dance_tempo",
              fluidRow(
                box(plotOutput("dance_tempo"), width = 12)
              )
      ),
      tabItem(tabName = "energy_tempo",
              fluidRow(
                box(plotOutput("energy_tempo"), width = 12)
              )
      ),
      tabItem(tabName = "acoustic_liveness",
              fluidRow(
                box(plotOutput("acoustic_liveness"), width = 12)
              )
              
      ),
      tabItem(tabName = "dance_energy",
              fluidRow(
                box(plotOutput("dance_energy"), width = 12)
              )
      ),
      
      tabItem(tabName = "dance_valance",
              fluidRow(
                box(plotOutput("dance_valance"), width = 12)
              )
      ),
      tabItem(tabName = "energy_valance",
              fluidRow(
                box(plotOutput("energy_valance"), width = 12)
              )
      )
    )
  )
)

server <- function(input, output) {
  output$genre <- renderPlot({
    spotify %>% 
      select(pop:musica_chihuahuense) %>% 
      gather(key = "genre", value = "count") %>% 
      ggplot(aes(x = reorder(genre, count), y = count)) +
      geom_bar(stat = "identity", fill = "springgreen3", color="black") +
      labs(title = "Number of Songs in Each Genre",
           x = "Genre",
           y = "Number of Songs")+
      coord_flip()+
      theme_minimal()+
      theme(text = element_text(size = 15))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$top5 <- renderPlot({
    spotify %>%
      count(artist_name) %>%
      top_n(6) %>%
      ggplot(aes(x = reorder(artist_name, -n), y = n)) +
      geom_bar(stat = "identity", fill = "springgreen3", color="black") +
      labs(title = "Top Artists with the Most Songs",
           x = NULL,
           y = "Number of Songs")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$pop <- renderTable({
    spotify %>% 
      filter(pop == 1) %>% 
      select(track_name, artist_name)
  })
  
  
  output$pop_genre <- renderPlot({
    spotify %>% 
      select(contains("pop")) %>% 
      select(-popularity) %>% 
      gather(key = "pop", value = "count") %>% 
      ggplot(aes(x = reorder(pop, count), y = count)) +
      geom_bar(stat = "identity", fill = "springgreen3", color="black") +
      labs(title = "All Pop Songs",
           x = NULL,
           y = "Number of Songs")+
      coord_flip()+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$canadian_pop <- renderTable({
    spotify %>% 
      filter(canadian_pop == 1) %>% 
      select(track_name, artist_name)
  })
  
  output$kpop <- renderTable({
    spotify %>%
      filter(k_pop > 0 | k_pop_girl_group >0) %>%
      mutate(k_pop_category = case_when(
        k_pop == 1 & k_pop_girl_group == 0 ~ "k-pop only",
        k_pop == 0 & k_pop_girl_group == 1 ~ "girl group only",
        k_pop == 1 & k_pop_girl_group == 1 ~ "both"
      )) %>%
      select(artist_name, track_name, k_pop_category)
  })
  
  output$duration <- renderPlot({
    spotify %>% 
      mutate(duration_ms = duration_ms/60000) %>%
      ggplot(aes(x = duration_ms)) +
      geom_density(fill = "springgreen3") +
      labs(title = " Average Duration of All Songs",
           x = "Duration (min)",
           y = "Density")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$loudness <- renderPlot({
    spotify %>%
      mutate(is_explicit = ifelse(is_explicit == 1, "Explicit", "Not Explicit")) %>%
      ggplot(aes(x = is_explicit, y = loudness, fill = is_explicit)) +
      geom_boxplot(fill="springgreen3") +
      labs(title = "Loudness of Explicit and Non-Explicit Songs",
           x = NULL,
           y = "Loudness")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$dance_tempo <- renderPlot({
    spotify %>%
      ggplot(aes(x = danceability, y = tempo)) +
      geom_point(size=2.5) +
      labs(title = "Analysis of the Danceability and Tempo of All Songs",
           x = "Danceability",
           y = "Tempo")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$energy_tempo <- renderPlot({
    spotify %>%
      ggplot(aes(x = energy, y = tempo)) +
      geom_point(size=2.5) +
      labs(title = "Analysis of the Energy and Tempo of All Songs",
           x = "Energy",
           y = "Tempo")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$acoustic_liveness <- renderPlot({
    spotify %>%
      ggplot(aes(x = acousticness, y = liveness)) +
      geom_point(size=2.5) +
      labs(title = "Analysis of the Acoustics and Liveness of All Songs",
           x = "Acousticness",
           y = "Liveness")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
    output$dance_energy <- renderPlot({
    spotify %>% 
      ggplot(aes(x = danceability, y = energy)) +
      geom_point(size = 2.5) +
      geom_smooth(method = "lm", color="springgreen3") +
      labs(title = "Analysis of the Danceability and Energy of All Songs",
           x = "Danceability",
           y = "Energy")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$dance_valance <- renderPlot({
    spotify %>%
      ggplot(aes(x = danceability, y = valence)) +
      geom_point(size=2.5) +
      geom_smooth(method = "lm", color="springgreen3") +
      labs(title = "Analysis of the Dancability and Valance of All Songs",
           x = "Danceability",
           y = "Valence")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
  output$energy_valance <- renderPlot({
    spotify %>%
      ggplot(aes(x = energy, y = valence)) +
      geom_point(size=2.5) +
      geom_smooth(method = "lm", color="springgreen3") +
      labs(title = "Analysis of the Energy and Valance of All Songs",
           x = "Energy",
           y = "Valence")+
      theme_minimal()+
      theme(text = element_text(size = 20))+
      theme(plot.title = element_text(hjust = 0.5))
  })
  
}

shinyApp(ui, server)
```


      
      
      
      
      
      
      

